---
title: "Libraries"
author: "Kabilan Kathiravel"
date: "05/12/2021"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, include = FALSE}
library(learnr)
library(tidyverse)
options(repos = "https://cran.stat.auckland.ac.nz/")
library(palmerpenguins)
```

## Welcome

This is a tutorial on libraries in R and the pipe operator. If you are not sure what a library is in this context, please check out [this link](https://hbctraining.github.io/Intro-to-R-flipped/lessons/04_introR_packages.html) to get some idea before trying the exercises. For the pipe operator try [here](https://style.tidyverse.org/pipes.html).

In this tutorial you will learn about:

- Loading and installing libraries
- Getting help to understand a specific library in R
- The pipe operator (%>%)

## Install packages

Let's have a look at the [Palmer Penguins dataset](https://allisonhorst.github.io/palmerpenguins/articles/intro.html). It contains size measurements for three penguin species observed on three islands in Antarctica's Palmer Archipelago. 

However, we don't already have the package installed in R so to use it we have to do so. Write down the command that installs the package ("palmerpenguins"), then press 'run code' to check your answer.

```{r installex1, exercise = TRUE}
    
```
```{r installex1-solution}
    install.packages("palmerpenguins")
```


<div id="installex1-hint">
**Hint:** This is the use of the `install.packages()` function.
</div>


## Load libraries and ask for help

To access the data, you need to load the "palmerpenguins" library into your R program. As it is now installed, you can do so!

```{r loadex1, exercise = TRUE}
    
```
```{r loadex1-solution}
    library(palmerpenguins)
```

You want to know more about the palmerpenguins library you are working with. There are two commonly used ways of doing this in R. Write down one below.

```{r helpex1, exercise = TRUE}
    
```
```{r helpex1-solution}
    help(palmerpenguins)
```

You need the tidyverse library to manipulate this data. You will also need ggplot2 to visualise it. Load each library in below.

```{r loadex2, exercise = TRUE}
    
```
```{r loadex2-solution}
    library(tidyverse)
    library(ggplot2)
```



## Piping

Now we have each library we need - excellent! Now we can get to looking at the data using the penguins dataset it is loaded to.

Let's have a look at the first six rows, using the pipe operator (%>%).

```{r pipeex1, exercise = TRUE}
    
```
```{r pipeex1-solution}
    penguins %>%
  head()
```

<div id="pipeex1-hint">
**Hint:** This is the use of the `head()` function.
</div>




## Visualisation

For visualisation we can use ggplot2. For example, we might want to see how flipper length varies with bill length. To do so, we can plot a scatter plot linking the two.

## Questions

```{r library-mcq, echo=FALSE}
question("Which of the following commands installs tidyverse in R?",
  answer("install(tidyverse)"),
  answer("library(tidyverse)"),
  answer("install.packages('tidyverse')", correct = TRUE),
  answer("open sesame")
)
```


